### 2021-04-12 input/proposal.tex
mostly on the start of adaptive scoped broadcast

Scoped broadcast is a communication primitive that propagates a
message around its broadcaster within an application-dependant scope.
For instance, a \process from Paris could scoped broadcast messages to
all \processes in Paris; and \processes from other cities would never
deliver such messages.
%% Dedicated scoped broadcast implementations
%% outperform uniform reliable broadcast
%% implementations~\cite{hadzilacos1994modular, raynal2013distributed}
%% since the former targets a connected \emph{subset} of the whole
%% network while the latter targets \emph{all} \processes of the
%% distributed system.
Scoped broadcast only targets a connected \emph{subset} of \processes
in the whole distributed system.
%
Distributed systems such as Edge infrastructures comprise
heterogeneous machines interconnected by communication
links. \Processes may crash but are not byzantine.

@@

In this section, we propose to dynamically adapt the scopes of \NAMEB
using \NAMEB itself for the sake of efficiency. \TODO{States of
  \processes change, predicates change over time.} At any time, a
\process decides whether or not it becomes the center (or
\emph{source}) of a scope (or \emph{partition}). \Processes must
eventually rally the partition corresponding to their closest source
despite concurrent operations, and dynamic changes in the network.

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy} = W_{yx}$ be the symmetric \underline{w}eight between
  $x$ and $y$, $\Pi_{xz}$ be the shortest \underline{p}ath from $x$ to
  $z$ the weight of which $|\Pi_{xz}|$ is lower than any other path
  weight, the only consistent partitioning $\mathcal{P}$ is a set of
  partitions $P_{s\in S}$ such that each \process belongs to a logical
  partition comprising its closest source: $\forall p \in P_{s_1},
  \nexists P_{s_2}$ such that $|\Pi_{s_1p}| > |\Pi_{s_2p}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of the
network state. For \processes to reach consistent partitioning
together, a \process becoming a source must notify every \process that
is closer to it than any other source, using \NAMEB for the sake of
efficiency.

\begin{theorem}[\underline{B}est \underline{e}ventual \underline{f}orwarding $(BEF) \iff CP$]
  Assuming reliable communication links where a correct \process $q$
  eventually receives the message $m$ \underline{s}ent to it by a
  \process $p$ ($s_{pq}(m) \implies r_{q}(m)$); a static system ($G^t =
  G^{t+1}$); assuming local knowledge only, \processes eventually reach
  consistent partitioning if and only if each \process eventually
  forwards its best known partition.
\end{theorem}

\TODO{If forwarding defined as send to all neighbors, then it also
  works in dynamic systems where \processes join and edges are added,
  for when a new neighbor appears, the \processes must forward their
  best.}

\begin{proof}
  \begin{asparadesc}
  \item [$BEF \implies CP$:] When a \process $s_1$ becomes a source, it
    belongs to its own partition, for there exists no better partition
    than its own: $\forall p \in V: |\Pi_{s_1 s_1}| < |\Pi_{s_1
      p}|$. It broadcasts such message to its neighbors. Since
    communication channels are reliable, neighboring \processes
    eventually receive such notification. In particular, whatever the
    order of received messages, every \process $p_1$ -- such that there
    exists no better partition $s_2$ than the received one -- delivers
    and forwards it: $\forall s_2 \in S: |\Pi_{p s_1}| < |\Pi_{p
      s_2}|$. The notification emanating from $s_1$ transitively
    reaches all such \processes through shortest paths. Since there
    exists only one best value per \process that can never be
    retracted, \processes eventually reach consistent
    partitioning. \TODO{Rework this with $s_{pq}$ and $r_{pq}$.}
  \item [$CP \implies BEF$:] \TODO{By contradiction, if there is a
    \process $p$ that does not forward, then another \process $q$ may
    never gets its best because it needs the weight $W_{pq}$.}
  \end{asparadesc}
\end{proof}

@@

Unfortunately, stale control information about deleted partitions may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about other deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issue with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending. Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ and \Process~$d$ do not deliver any $\alpha_d$ or
$\delta_d$ since they already delivered it. The system converges to an
unconsistent state where some \processes assume they belong to a
partition while it does not exist anymore. Naive propagation of
$\alpha$ and $\delta$ messages is insufficient to guarantee consistent
partitioning when each \process can add or delete its partition at any
time.

>>>>

Scoped broadcast is a communication primitive that propagates a
message around its broadcaster within an application-dependant scope.
For instance, a \process from Paris could scoped broadcast messages to
all \processes in Paris; and \processes from other cities would never
deliver such messages.
Scoped broadcast only targets a connected \emph{subset} of \processes
in the whole distributed system.
%
Distributed systems such as Edge infrastructures comprise
heterogeneous machines interconnected by communication
links. \Processes may crash but are not byzantine.

@@

In this section, we propose to dynamically adapt the scopes of \NAMEB
using \NAMEB itself for the sake of efficiency: a \NAMEB message may
modify the state of a \process, consequently changing predicate
evaluations over deliveries.  At any time, a \process decides whether
or not it becomes the center (or \emph{source}) of a scope (or
\emph{partition}). \Processes must eventually rally the partition
corresponding to their closest source despite concurrent operations,
and dynamic changes in the network.

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy} = W_{yx}$ be the positive symmetric \underline{w}eight
  between $x$ and $y$, $\Pi_{xz}$ be the shortest \underline{p}ath
  from $x$ to $z$ the weight of which $|\Pi_{xz}|$ is lower than any
  other path weight, the only consistent partitioning $\mathcal{P}$ is
  a set of partitions $P_{s\in S}$ such that each \process belongs to
  a logical partition comprising its closest source: $\forall p \in
  P_{s_1}, \nexists P_{s_2}$ such that $|\Pi_{s_1p}| > |\Pi_{s_2p}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \process becoming a source must notify every \process that
is closer to it than any other source. As stated in
Section~\ref{sec:scoped}, since epidemic dissemination and scoped
broadcast have well-aligned preoccupations, we assume implementations
relying on forwarding of messages from neighbor to neighbor.

\begin{theorem}[\underline{B}est \underline{e}ventual \underline{f}orwarding $(BEF) \iff CP$]
  Assuming reliable communication links where a correct \process $q$
  eventually receives the message $m$ \underline{s}ent to it by a
  \process $p$ ($s_{pq}(m) \implies r_{q}(m)$); a monotonically
  growing system ($G^t \subseteq G^{t+1}$); \processes eventually
  reach consistent partitioning if and only if each \process
  eventually forwards its best known partition.
\end{theorem}

\begin{proof}
  \begin{asparadesc}
  \item [$BEF \implies CP$:] When a \process $s_1$ becomes a source,
    it belongs to its own partition, for there exists no better
    partition than its own: $\forall p \in V: |\Pi_{s_1 s_1}| <
    |\Pi_{s_1 p}|$. It forwards such message to its neighbors. Since
    communication links are reliable, neighboring \processes
    eventually receive such notification. Most importantly, whatever
    the order of received messages, every \process $p_1$ -- such that
    there exists no better partition $s_2$ than the received one --
    delivers and forwards it: $\forall s_2 \in S: |\Pi_{p s_1}| <
    |\Pi_{p s_2}|$. The message originating from $s_1$ transitively
    reaches all such \processes through shortest paths. Since there
    exists only one best value per \process that can never be
    retracted, \processes eventually reach consistent
    partitioning. % \TODO{Rework this with $s_{pq}$ and $r_{pq}$.}
  \item [$CP \implies BEF$:] \TODO{By contradiction, if there is a
    \process $p$ that does not forward, then another \process $q$ may
    never gets its best because it needs the weight $W_{pq}$.}
  \end{asparadesc}
\end{proof}

@@

Unfortunately, stale control information about deleted partitions may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about other deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issue with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending. Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ and \Process~$d$ do not deliver any $\alpha_d$ or
$\delta_d$ since they already delivered it. The system converges to an
inconsistent state where some \processes assume they belong to a
partition while it does not exist anymore. Naive propagation of
$\alpha$ and $\delta$ messages is insufficient to guarantee consistent
partitioning when each \process can add or delete its partition at any
time.
